---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)
library(pROC)
library(modelr)
```

```{r}
lending <- read_csv("Final_Project/clean data/leanding_cleaning.csv")
```

```{r}
lending_model <- lending %>% 
  select(-c(loan_amnt, installment, emp_title, issue_d, desc, title, state_abb, total_pymnt_inv, profit_loss, state_name, sub_grade)) %>% 
  mutate_if(is_character, as_factor)
```

```{r}
alias(loan_status ~ ., data = lending_model)
```


```{r}
lending_model1 <- lending_model %>% 
  select(term, int_rate, emp_length, home_ownership, interest_rate, loan_status)
```

```{r}
lending_model2 <- lending_model %>% 
  select(annual_inc, verification_status, purpose, loan_status)
```

```{r}
lending_model3 <- lending_model %>% 
select(dti, delinq_2yrs, fico_range_low, total_acc, grade, loan_status)
```

```{r message=FALSE, warning=FALSE}
ggpairs(lending_model1)
```

```{r message=FALSE, warning=FALSE}
ggpairs(lending_model2)
```

```{r message=FALSE, warning=FALSE}
ggpairs(lending_model3)
```

```{r}
model_a <- glm(loan_status ~ .,
               data = lending_model1,
               family = binomial(link = "logit"))
summary(model_a)
```

From model a the most important factors are interest rate (numeric version) and term of loan. 

```{r}
model_b <- glm(loan_status ~ ., 
               data = lending_model2,
               family = binomial(link = "logit"))

summary(model_b)
```
**use anova for testing significance of purpose**

From model b the most important factors are annual income and verification status. The purpose for the loan is mixed so should be explored. 

**verification_status taken out because it no longer fits a good model**

```{r}
model_c <- glm(loan_status ~ .,
               data = lending_model3,
               family = binomial(link = "logit"))
summary(model_c)
```

From model c the most important factors are grade, total_acc and dti. I would also like to keep fico as it is the equivalent of a credit score. 

**total_acc was taken out of model as when added with other financial categories it was no longer significant.**

```{r}
model_finance <- glm(loan_status ~ annual_inc + dti + fico_range_low,
                     data = lending_model,
                     family = binomial(link = "logit"))

summary(model_finance)
```

```{r}
model_category <- glm(loan_status ~ term + int_rate,
                      data = lending_model,
                      family = binomial(link = "logit"))

summary(model_category)
```


```{r}
model_grade <- glm(loan_status ~ sub_grade,
                   data = lending_model,
                   family = binomial(link = "logit"))

summary(model_grade)
```

```{r}
lending_data_finance <- lending_model %>% 
  add_predictions(model_finance, type = "response")

head(lending_data_finance)
```

```{r}
lending_data_category <- lending_model %>% 
  add_predictions(model_category, type = "response")
```

```{r}
lending_data_grades <- lending_model %>% 
  add_predictions(model_grade, type = "response")
```

```{r}
roc_lend_fin <- lending_data_finance %>% 
  roc(response = loan_status, predictor = pred)

roc_lend_cat <- lending_data_category %>% 
  roc(response = loan_status, predictor = pred)

roc_lend_grade <- lending_data_grades %>% 
  roc(response = loan_status, predictor = pred)
```

```{r}
roc_curve <- ggroc(
  data = list(
    mod1 = roc_lend_fin,
    mod2 = roc_lend_cat,
    mod3 = roc_lend_grade
  ),
  legacy.axes = TRUE) +
  coord_fixed()

roc_curve
```

```{r}
auc(roc_lend_fin)
auc(roc_lend_cat)
auc(roc_lend_grade)
```

**Back to the drawing board** 

Let's see how we get on with a potentially overfit model.


```{r}
model_all <- glm(loan_status ~ annual_inc + dti + fico_range_low + term + int_rate,
                 data = lending_model,
                 family = binomial(link = "logit"))
summary(model_all)
```

```{r}
lend_data_all <- lending_model %>% 
  add_predictions(model_all, type = "response")
```

```{r}
roc_lend_all <- lend_data_all %>% 
  roc(response = loan_status, predictor = pred)
```

```{r}
roc_curve_all <- ggroc(data = list(
  "Built Model" = roc_lend_all,
  "Grade Model" = roc_lend_grade),
                   legacy.axes = TRUE) +
  labs(x = "Specificity", y = "Sensitivity", linetype="Model Name")+ 
  coord_fixed()

roc_curve_all
```

```{r}
auc(roc_lend_all)
auc(roc_lend_grade)
```

```{r}
library(caret)
```

```{r}
cv_10_fold_model <- trainControl(method = "cv",
                                 number = 10,
                                 savePredictions = TRUE)
```

```{r}
model <- train(loan_status ~ annual_inc + dti + fico_range_low + term + int_rate,
               data = lending_model,
               trControl = cv_10_fold_model,
               method = "glm",
               family = binomial(link = "logit"),
               metric = "accuracy")
```

```{r}
summary(model)
```

```{r}
model$pred
```

```{r}
model$resample
```

```{r}
mean(model$resample$Kappa)
```

**new k-fold model**

```{r}
folds <- createFolds(lending_model$loan_status, k = 10, list = TRUE, returnTrain = FALSE)

lr_auc <- function(train, test) {
  lr_model <- glm(loan_status ~ annual_inc + dti + fico_range_low + term + int_rate, data = train, family = binomial)
  pred <- predict(lr_model, test, type = "response")
  roc(test$loan_status, pred)$auc
}

```

```{r}
lending_model_test1 <- lending_model %>% 
  slice(folds$Fold01)

lending_model_train1 <- lending_model %>% 
  slice(-folds$Fold01)

lr_auc_results1 <- lr_auc(lending_model_train1, lending_model_test1)
```
```{r}
auc_results <- NULL
```

```{r}
for (i in 1:length(folds)) {
  auc_results[i] <- lr_auc(lending_model %>% slice(-folds[[i]]),
                           lending_model %>% slice(folds[[i]]))
}

mean(auc_results)
```

```{r}
lr_auc_grades <- function(train, test) {
  lr_model_grades <- glm(loan_status ~ grade, data = train, family = binomial)
  pred <- predict(lr_model_grades, test, type = "response")
  roc(test$loan_status, pred)$auc
}
```


```{r}
auc_results_grades <- NULL
for (i in 1:length(folds)) {
  auc_results_grades[i] <- lr_auc_grades(lending_model %>% slice(-folds[[i]]),
                           lending_model %>% slice(folds[[i]]))
}
  
mean(auc_results_grades)
```
```{r}

```

