---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

```{r}
lending <- read_csv("../clean data/leanding_cleaning.csv")
```

```{r}
lending %>% 
  ggplot(aes(x = issue_d, y = profit_loss))+
  geom_col()
```
```{r}
ggplot(lending) +
  geom_bar(aes(x = purpose, fill = loan_status), position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = purpose, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = sub_grade, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = grade, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = state_name, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = home_ownership, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = verification_status, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = interest_rate, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = emp_length, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = term, fill = loan_status))+
  geom_bar(position = "dodge")+
  coord_flip()
```

```{r}
lending %>% 
  group_by(grade) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  group_by(interest_rate) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  ggplot(aes(x = interest_rate, y = profit_loss))+
  geom_col()
```

```{r}
lending %>% 
  ggplot(aes(x = int_rate, y = profit_loss))+
  geom_point()
```

```{r}
lending %>% 
  group_by(purpose) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  ggplot(aes(x = purpose, y = profit_loss))+
  geom_col()+
  coord_flip()
```

```{r}
lending %>% 
  ggplot(aes(x = annual_inc, y = profit_loss))+
  geom_point()
```

```{r}
lending %>% 
  group_by(loan_status) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  group_by(sub_grade) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  group_by(term) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  ggplot(aes(x = loan_amnt, y = profit_loss))+
  geom_point()
```

```{r}
lending %>% 
  ggplot(aes(x = sub_grade, y = annual_inc))+
  geom_col()
```

```{r}
lending %>% 
  group_by(grade) %>% 
  summarise(int_rate = mean(int_rate)) %>% 
  ggplot(aes(x = grade, y = int_rate, fill = grade)) +
  geom_col()+
  scale_fill_manual(values=c("goldenrod", "#9933FF", "#33FFFF", "red", "darkblue", "darkolivegreen4", "dodgerblue"))+
   labs(x = "Grade\n", y = "\nInterest Rate (%)", title = "Interest Rate over LendingClub Grades\n")+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  coord_flip() 

ggsave("gradesvsinterest.jpg")

```

```{r}
lending %>% 
  group_by(grade) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  mutate(profit_loss = profit_loss/1000) %>% 
  ggplot(aes(x = grade, y = profit_loss, fill = grade))+
  geom_col()+ 
  scale_fill_manual(values=c("goldenrod", "#9933FF", "#33FFFF", "red", "darkblue", "darkolivegreen4", "dodgerblue"))+
   labs(x = "Grade\n", y = "\nProfitability($US1000)", title = "Profitability over Grade\n")+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  coord_flip() 

ggsave("profitovergrade.jpg")
```

```{r}
lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  arrange(profit_loss) %>% 
  head(10) %>% 
  ggplot(aes(x = state_name, y = profit_loss))+
  geom_col()+
  coord_flip()
```

```{r}
lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  arrange(desc(profit_loss)) %>% 
  head(10) %>% 
  ggplot(aes(x = state_name, y = profit_loss))+
  geom_col()+
  coord_flip()
```

```{r}
lending %>% 
  summarise(loan_amnt = mean(loan_amnt))
```

```{r}
lending %>% 
  summarise(max_loan = max(loan_amnt))
```

```{r}
lending %>% 
  summarise(min_loan = min(loan_amnt))
```


```{r}
lending %>% 
  summarise(loan_amnt = median(loan_amnt))
```

```{r}
lending %>% 
  summarise(int_rate = mean(int_rate))
```
```{r}
lending %>% 
  summarise(max_int = max(int_rate))
```

```{r}
lending %>% 
  summarise(min_int = min(int_rate))
```

```{r}
lending %>% 
  group_by(emp_length) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  group_by(home_ownership) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  group_by(emp_length) %>% 
  summarise(annual_inc = mean(annual_inc))
```

```{r}
lending %>% 
  group_by(emp_length) %>% 
  summarise(int_rate = mean(int_rate))
```

```{r}
lending %>% 
  group_by(issue_d) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  ggplot(aes(x = fico_range_low, y = profit_loss))+
  geom_point()
```

```{r}
lending %>% 
  group_by(verification_status) %>% 
  summarise(profit_loss = sum(profit_loss))
```

```{r}
lending %>% 
  group_by(sub_grade) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  ggplot(aes(x = sub_grade, y = profit_loss))+
  geom_point()+
  geom_line(group = 1)
```

```{r}
lending %>% 
  ggplot(aes(x = loan_status))+
  geom_histogram(stat = "count")
```

```{r}
lending %>% 
  group_by(loan_status) %>% 
  count(loan_status)
```

```{r}
default_freq <- table(lending$loan_status)

default_prop <- prop.table(default_freq)

default_prop
```

```{r}
lending %>% 
  group_by(sub_grade) %>% 
  summarise(int_rate = mean(int_rate)) %>% 
  ggplot(aes(x = sub_grade, y = int_rate, fill = sub_grade))+
  geom_col()+
  scale_fill_manual(values=c("goldenrod", "goldenrod", "goldenrod", "goldenrod", "goldenrod",
                             "#9933FF", "#9933FF", "#9933FF", "#9933FF", "#9933FF",
                            "#33FFFF", "#33FFFF", "#33FFFF", "#33FFFF", "#33FFFF",
                             "red", "red", "red", "red", "red",
                             "darkblue",  "darkblue",  "darkblue",  "darkblue",  "darkblue",
                            "darkolivegreen4", "darkolivegreen4", "darkolivegreen4", "darkolivegreen4", "darkolivegreen4",
                            "dodgerblue", "dodgerblue", "dodgerblue", "dodgerblue", "dodgerblue"))+
  labs(x = "Grade", y = "Interest Rate (%)", title = "Interest Rate over LendingClub Grades")+
  theme(legend.position = "none")+
  coord_flip() 

ggsave("LC_grading.pdf")
ggsave("LC_grading.jpg")

```

```{r}
lending %>% 
  filter(state_name == "Nevada") %>% 
  summarise(annual_inc = mean(annual_inc))
```
```{r}
lending %>% 
  filter(state_name == "Nevada") %>% 
  group_by(purpose) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  ggplot(aes(x = purpose, y = profit_loss))+
  geom_col()+
  coord_flip()
```

```{r}
nevada <- lending %>% 
  filter(state_name == "Nevada")

nevada_default_freq <- table(nevada$loan_status)

nevada_default_prop <- prop.table(nevada_default_freq)

nevada_default_prop
```


```{r}
lending %>% 
  filter(state_name == "New York") %>% 
  summarise(annual_inc = mean(annual_inc))
```

```{r}
lending %>% 
  group_by(purpose) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  mutate(profit_loss = profit_loss/1000) %>% 
  ggplot(aes(x = reorder(purpose, profit_loss), y = profit_loss, fill = profit_loss < 0))+
  geom_col()+
  labs(x = "Purpose for Loan\n", y = "\nProfit/Loss(US$1000s)", title = "Profitability of Loan Purpose\n")+
   scale_fill_manual(guide = FALSE,
                          name = 'change < 0', 
                          values = setNames(c('indianred', 'dodgerblue'), c(T, F)))+
  theme_minimal()+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  coord_flip()

ggsave("profitabilitypurpose.jpg")
```

```{r}
new_york <- lending %>% 
  filter(state_name == "New York")

new_york_default_freq <- table(new_york$loan_status)

new_york_default_prop <- prop.table(new_york_default_freq)

new_york_default_prop
```

```{r}
new_york %>% 
  summarise(loan_amnt = mean(loan_amnt),
dti = mean(dti),
fico_range_low = mean(fico_range_low),
int_rate = mean(int_rate))
```
```{r}
nevada %>% 
  summarise(loan_amnt = mean(loan_amnt),
dti = mean(dti),
fico_range_low = mean(fico_range_low),
int_rate = mean(int_rate))
```

```{r}
lending %>% 
  filter(state_name == "New York" | state_name == "Nevada") %>% 
  group_by(purpose, state_name) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  ggplot(aes(x = reorder(purpose, profit_loss), y = profit_loss, fill = profit_loss < 0))+
  geom_col()+
  facet_wrap(~state_name, scales = "free_y")+
    labs(x = "Purpose for Loan", y = "Profit/Loss(US$)", title = "State Difference in Profitability of Loan Purpose")+
  theme(axis.text.x = element_text(angle=45, 
                                       hjust = 1))+
  theme_minimal()+
   theme(axis.text.x = element_text(angle=45, 
                                       hjust = 1))+
   scale_fill_manual(guide = FALSE,
                          name = 'change < 0', 
                          values = setNames(c('indianred', 'dodgerblue'), c(T, F)))

ggsave("Nevada_New_York.jpg")
```

```{r}
loss_states <- lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  arrange(profit_loss) %>% 
  head(5)

loss_states

high_states <- lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  arrange(desc(profit_loss)) %>% 
  head(5)

money_states <- bind_rows(loss_states, high_states) %>% 
  arrange(desc(profit_loss))

money_states %>% 
  mutate(profit_loss = profit_loss/1000) %>% 
  ggplot(aes(x = reorder(state_name, profit_loss), y = profit_loss, fill = profit_loss < 0))+
  geom_bar(stat = "identity")+
    labs(x = "US State\n", y = "\nProfit/Loss(US$1000s)", title = "Most and Least Profitable US States\n")+
  theme_minimal()+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  scale_fill_manual(guide = FALSE,
                          name = 'change < 0', 
                          values = setNames(c('indianred', 'dodgerblue'), c(T, F)))+
  coord_flip()

ggsave("ProfitableStates.jpg")
                         
```

```{r}

loss_states <- lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = sum(profit_loss)) %>% 
  arrange(profit_loss) %>% 
  head(5)

loss_states %>% 
  mutate(profit_loss = profit_loss / 1000) %>% 
  ggplot(aes(x = reorder(state_name, profit_loss), y = profit_loss, fill = profit_loss < 0)) +
  geom_bar(stat = "identity") +
  labs(x = "US State\n", y = "\nTotal Loss per State (US$ in Thousands)", title = "Least Profitable US States\n") +
  theme_minimal() +
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16)) +
  scale_fill_manual(guide = FALSE,
                    name = 'Loss > 0', 
                    values = setNames(c('indianred'), c(T))) +
  coord_flip()


ggsave("LeastProfitableStates.jpg")
                         
```
```


```{r}
lending %>% 
  group_by(sub_grade) %>% 
  count(sub_grade)
  summarise(sub_grade_median = median(n))
```

```{r}
lending %>% 
  group_by(purpose) %>% 
  summarise(profit_loss = mean(profit_loss)) %>% 
  ggplot(aes(x = reorder(purpose, profit_loss), y = profit_loss, fill = profit_loss < 0))+
  geom_col()+
  labs(x = "Purpose for Loan\n", y = "\nProfit/Loss(US$)", title = "Profitability of Loan Purpose\n")+
   scale_fill_manual(guide = FALSE,
                          name = 'change < 0', 
                          values = setNames(c('indianred', 'dodgerblue'), c(T, F)))+
  theme_minimal()+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  coord_flip()

ggsave("meanprofitabilitypurpose.jpg")
```


```{r}
loss_states2 <- lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = mean(profit_loss)) %>% 
  arrange(profit_loss) %>% 
  head(5)

loss_states2

high_states2 <- lending %>% 
  group_by(state_name) %>% 
  summarise(profit_loss = mean(profit_loss)) %>% 
  arrange(desc(profit_loss)) %>% 
  head(5)

money_states2 <- bind_rows(loss_states2, high_states2) %>% 
  arrange(desc(profit_loss))

money_states2 %>% 
  ggplot(aes(x = reorder(state_name, profit_loss), y = profit_loss, fill = profit_loss < 0))+
  geom_bar(stat = "identity")+
    labs(x = "US State\n", y = "\nProfit/Loss(US$)", title = "Most and Least Profitable US States\n")+
  theme_minimal()+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  scale_fill_manual(guide = FALSE,
                          name = 'change < 0', 
                          values = setNames(c('indianred', 'dodgerblue'), c(T, F)))+
  coord_flip()

ggsave("meanprofitabilitystate.jpg")

```

```{r}
lending %>% 
  group_by(grade) %>% 
  summarise(profit_loss = mean(profit_loss)) %>% 
  ggplot(aes(x = grade, y = profit_loss, fill = grade))+
  geom_col()+ 
  scale_fill_manual(values=c("goldenrod", "#9933FF", "#33FFFF", "red", "darkblue", "darkolivegreen4", "dodgerblue"))+
   labs(x = "Grade\n", y = "\nProfitability($US)", title = "Profitability over Grade\n")+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  coord_flip() 

ggsave("meanprofitabilitygrade.jpg")

```

```{r}
lending %>% 
  group_by(grade) %>% 
  summarise(loan_amnt = sum(loan_amnt)) %>% 
  ggplot(aes(x = grade, y = loan_amnt, fill = grade))+
  geom_col()+ 
  scale_fill_manual(values=c("goldenrod", "#9933FF", "#33FFFF", "red", "darkblue", "darkolivegreen4", "dodgerblue")) +
   theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))+
  coord_flip() 
```

